%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = Thesis_Driver.tex

\chapter{Acknowledgments}
I would like to first thank my supervisor, Professor Doug Oldenburg, for giving me the opportunity to join the UBC-GIF group and for his tireless excitement for research. Your passion for science is contagious.

I want to acknowledge the outstanding contribution of many colleagues from the open-source community. Thanks to Rowan Cockett, Lindsey Heagy and Seogi Kang for establishing the SimPEG framework and pushing me to learn Python. Special mention goes to Joe Capriotti for his key update to the Octree mesh discretization. The work presented in this thesis would not have been possible without your contribution. Thank you to the rest of the UBC-GIF group for your support, kind words and friendship.

I am grateful to the companies and staff that have supported research at UBC-GIF over all these years: Anglo American, Barrick, BHP, Cameco, CGI, Glencore, Newmont, Rio Tinto, Teck, Vale. Personal thanks to John McGaughey at Mira Geosciences for keeping on staff over all these years, as well as for lending me valuable hardware and software.

Thank you to the many public libraries across Canada where this thesis was written. Cities and towns from east to west (*with great pool nearby):
Ottawa, Winnipeg*, Prince George, Fort Saint-James, Smithers*, Vancouver, Richmond*, Steveston, Nanaimo, Courtney, Alert Bay, Port Hardy.

Last but not least, extra special thanks goes to my best friend, my partner in crimes and lover Chantelle Bellrichard. I am done being a student now. Thank you to \emph{sister-wife} Candice Nichol for providing distractions to \emph{first-wife}, allowing me to finish my thesis. So many more BBQ \& sunsets at the beach to be enjoyed ...


